### 2020-11-08 meeting-notes/20201108-expose-input-language.tex
first step at typed interaction

\documentclass{article}

\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{mathpartir}
\usepackage{listings}
\usepackage{url}

\usepackage{hyperref}

\usepackage{xcolor}
\usepackage{tcolorbox}
\tcbuselibrary{breakable}

\newenvironment{changethis}{%
  \begin{tcolorbox}[breakable,notitle,boxrule=0pt,colback=blue!20,colframe=blue!20]}{%
  \end{tcolorbox}}

\newenvironment{ethereum}{%
  \begin{tcolorbox}[breakable,notitle,boxrule=0pt,colback=red!20,colframe=red!20]}{%
  \end{tcolorbox}}

\input{solidity-highlighting}

\title{Expos\'{e}\\
  Input Language for Smart Contract API}
\author{Peter Thiemann}
\begin{document}
\maketitle{}

\section{Introduction}
\label{sec:introduction}

The current Tezos API facing to OCaml (as well as the one provided by
the Tezos client) is untyped. Here we propose a DSL to make it typed.
First we recall some information from the Michelson white paper
\url{https://tezos.gitlab.io/whitedoc/michelson.html#core-data-types-and-notations}.

\subsection{Core data types and notations}

\begin{itemize}
\item \texttt{string}, \texttt{nat}, \texttt{int} and \texttt{bytes}: The core primitive constant types.
\item 
  \texttt{bool}: The type for booleans whose values are \texttt{True} and \texttt{False}.
\item 
  \texttt{unit}: The type whose only value is \texttt{Unit}, to use as a placeholder
  when some result or parameter is not necessary. For instance, when
  the only goal of a contract is to update its storage.
\item 
  \texttt{list (t)}: A single, immutable, homogeneous linked list, whose
  elements are of type (t), and that we write \texttt{\{\}} for the empty list or
  \texttt{\{ first ; ... \}}. 
\item 
  \texttt{pair (l) (r)}: A pair of values a and b of types (l) and (r), that we
  write \texttt{(Pair a b)}.
\item 
  \texttt{option (t)}: Optional value of type (t) that we write \texttt{None} or \texttt{(Some
  v)}.
\item 
  \texttt{or (l) (r)}: A union of two types: a value holding either a value a
  of type (l) or a value b of type (r), that we write \texttt{(Left a)} or
  \texttt{(Right b)}.
\item 
  \texttt{set (t)}: Immutable sets of values of type (t) that we write as lists
  \texttt{\{ item ; ... \}}, of course with their elements unique, and sorted.
\item 
  \texttt{map (k) (t)}: Immutable maps from keys of type (k) of values of type
  (t) that we write \texttt{\{ Elt key value ; ... \}}, with keys sorted.
\item 
  \texttt{big\_map (k) (t)}: Lazily deserialized maps from keys of type (k) of
  values of type (t) that we write \texttt{\{ Elt key value ; ... \}}, with keys
  sorted. These maps should be used if one intends to store large
  amounts of data in a map. They have higher gas costs than standard
  maps as data is lazily deserialized. A \texttt{big\_map} cannot appear inside
  another \texttt{big\_map}.
\end{itemize}

\subsection{Relation to OCaml datatypes}
\label{sec:relat-ocaml-datatyp}

\begin{center}
  \begin{tabular}{|l|l|}
    Michelson & OCaml \\\hline
    \texttt{string}& \texttt{string}  \\
    \texttt{nat} & ? (\texttt{int}) \\
    \texttt{int} & \texttt{int} \\ 
    \texttt{bytes}& ? \\
    \texttt{bool}&\texttt{bool} \\
    \texttt{unit}&\texttt{unit} \\
    \texttt{list t}&\texttt{t list} \\
    \texttt{pair l r}&\texttt{l * r} \\
    \texttt{option t}     &\texttt{t option} \\
    \texttt{or l r}&\texttt{[`Left of l | `Right of r]} \\
    \texttt{set t}&? \texttt{Set.Make (T).t} \\
    \texttt{map k t}&? \texttt{`t Map.Make(K).t} \\
    \texttt{big\_map k t}&? \texttt{`t Map.Make(K).t} \\
    \hline
  \end{tabular}
\end{center}

\section{Tasks}
\label{sec:tasks}

We can represent Michelson parameter values with the following OCaml type using GADTs.
\begin{lstlisting}[language=Caml]
type _ Mvalue =
  | Mstring : string -> string Mvalue
  | Mint  : int -> int Mvalue
  | Mbool : bool -> bool Mvalue
  | Munit : unit -> unit Mvalue
  | Mnil  : 't list Mvalue
  | Mcons : 't Mvalue * 't list Mvalue -> 't list Mvalue
  | Mpair : 'l Mvalue * 'r Mvalue -> ('l * 'r) Mvalue
  | Mnone : 't option Mvalue
  | Msome : 't Mvalue -> 't option Mvalue
  | Mleft : 'l Mvalue -> [`Left of 'l | `Right 'r] Mvalue
  | Mright : 'r Mvalue -> [`Left of 'l | `Right 'r] Mvalue
\end{lstlisting}

\begin{lstlisting}[language=Caml]
type _ Mtype = 
  | Tstring : string Mtype
  | Tint    : int Mtype
  | Tbool   : bool Mtype
  | Tunit   : unit Mtype
  | Tlist   : 't Mtype -> 't list Mtype
  | Tpair   : 'l Mtype * 'r Mtype -> ('l * 'r) Mtype
  | Toption : 't Mtype -> 't option Mtype
  | Tor     : 'l Mtype * 'r Mtype -> [`Left of 'l | `Right of 'r] Mtype
\end{lstlisting}
Instead of using polymorphic variants, one could also define a binary
sum type like Haskell's Either type.
\begin{lstlisting}[language=Caml]
type ('a, 'b) either = Left of 'a | Right of 'b
\end{lstlisting}

To read input values, we'd need a function like the following.
\begin{lstlisting}[language=Caml]
mread : 't Mtype -> 't -> 't Mvalue
\end{lstlisting}

Or given a contract for index type \lstinline/t/, we'd have an
invocation function
\begin{lstlisting}[language=Caml]
invoke : 't Mtype Tezos.contract -> 't -> result Tezos.apimonad
\end{lstlisting}
with an internal serialization function
\begin{lstlisting}[language=Caml]
serialize : 't Mtype -> 't -> string
\end{lstlisting}
Not sure if deserialization is also needed, but it might be necessary
to decode the current state of a contract. Deserialization may fail
and thus raise an exception.
\begin{lstlisting}[language=Caml]
deserialize : 't Mtype -> string -> 't
\end{lstlisting}
\end{document}