### 2021-01-18 papers/execution-model/model.tex
changed puh calculation and storing of contract

\documentclass[a4paper]{llncs}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage{mathpartir}
\usepackage{fixme}

\input{macros}

\title{Blockchain System}

\author{}
\institute{}

\date{\today}

\begin{document}
\maketitle
\pagestyle{plain}

\section{Definitions}
\begin{definition}%[Set]
Let $\mathbf{Elt}$ be the set of (concrete) elements. Let $\emptyset$ be an empty set and $\mathbf{e}$ $\in$  $\mathbf{Elt}$. A set of elements is expressed as the following syntax:
\\
$\mathbf{s} :: = \emptyset \mid \mathbf{e} \mid \mathbf{s :: s}$
\end{definition}

\subsection{Local Node}
\label{sec:local-node}

\begin{definition}%[Account]
  An implicit \emph{account} is represented by a key pair $\langle
  %\ALS,
  \PAK, \PUK
  %, \PKH
  \rangle $, where
  $\PAK$ is its private key and $\PUK$ is its public key.
\end{definition}

\begin{definition}%[Operation]
  \emph{Operations} are defined by the following grammar:
  \begin{align*}
    \OP &::= \TRANSFER[\PARAMETER]\NTEZ\PUK\ADDR\MTEZ
    \\&\mid \ORIGINATE\NTEZ\PUK\CODE\INIT\MTEZ
  \end{align*}
  where \NTEZ\ and \MTEZ\ are tokens, \ADDR\ is either a public
  key for an implicit account  or a public hash for a smart contract,
  \PARAMETER\ is an argument passed to the smart contract's script,
  which is empty if it is a transfer to an implicit account, \CODE\ is
  a script of the smart contract, \INIT\ is an initial value of the
  contract's storage. 
\end{definition}

\begin{definition}%[Query]
\emph{Queries} are defined by the following grammar:
\begin{align*}
  \QRY & ::=
  \GETCOUNTER\PUK  
  \\&\mid \GETBALANCE\ADDR
  \\&\mid \GETSTATUS\OPH
  \\&\mid \GETSTORAGE\PUH
  \\&\mid \GETCODE\PUH 
  \\&\mid \GETTYPE\PUH
\end{align*}
\end{definition}

\begin{definition}%[Program]
\emph{Expressions} are defined by the following grammar:
\begin{align*}
  \CONSTANT & ::= \INT %\mid \ADDR
              \mid \OPH \mid \PUH \mid\PUK \mid \CODE \mid
              \NTEZ \mid \SUNIT \mid \FALSE \mid \TRUE \\
  \STATUS & ::= \STATUSPENDING \mid \STATUSINCLUDING (\INT) \mid
            \STATUSTIMEOUT \\
  \ERROR &::= \ERRPRG \mid \ERRBAL \mid \ERRCOUNT \mid \ERRFEE \mid
           \ERRPUK \mid \ERRPUH \mid \ERRARG \mid \ERRINIT\\
	\EXPR & :: =  \CONSTANT \mid \STATUS \mid \ERROR \mid \VARIABLE \mid \lambda \VARIABLE. \EXPR\ 
	\mid \EXPR\EXPR \mid (\EXPR,\EXPR) \mid \FST\ \EXPR \mid \SND\ \EXPR 	%\mid \NI \EI
	\\&\mid \EXPR\ \PLUS\ \EXPR\ \mid \EXPR\ \MINUS\ \EXPR\ \mid \EXPR\ \EQUAL\ \EXPR\ \mid \EXPR\ \LESS\ \EXPR
	\\&\mid \EXPR\ \AND\ \EXPR\ \mid \EXPR\ \OR\ \EXPR\ \mid \NOT\ \EXPR
  \\&\mid \NIL\ \mid \CONS\ \EXPR\ \EXPR \mid \LEFT\ \EXPR \mid\RIGHT\
  \EXPR \mid \SOME\ \EXPR \mid \NONE
  \\& \mid \RAISE\ \EXPR \mid \TRY\ \EXPR\ \EXCEPT\ \EXPR
  \\& \mid \MATCH\ \EXPR\ \WITH\ \PATTERN\to\EXPR \dots
  \\& \mid \TRANSFER[\EXPR]\EXPR\EXPR\EXPR\EXPR
  \\& \mid \ORIGINATE\EXPR\EXPR\EXPR\EXPR\EXPR
  \\&\mid \QOP\  \EXPR\\
  \QOP & ::= balance \mid status \mid storage \mid code \mid counter \mid type
\end{align*}
We let $\EXPRS$ range over sets of expressions.

We let $\ECS$ range over evaluation contexts. They are defined as
follows.
\begin{align*}
  \EC{} & :: = \SC[\overline\VAL\ \EC{}\ \overline\EXPR] \mid \RAISE\ \EC{} \mid \TRY\ \EC{}\ \EXCEPT\ \EXPR
  \mid \MATCH\ \EC{}\ \WITH\ \PATTERN\to\EXPR \dots
\end{align*}
Here $\SC$ ranges over the remaining syntactic constructors, which are
treated uniformly: evaluation proceeds from left to right.
\begin{align*}
  \VAL & ::= \CONSTANT \mid \STATUS \mid \ERROR \mid \lambda x.\EXPR \mid(\VAL, \VAL) \mid
         \NIL \mid \CONS\ \VAL\ \VAL \mid \LEFT\ \VAL \mid \RIGHT\
         \VAL \mid \SOME\ \VAL\mid \NONE
\end{align*}
\end{definition}

Example pseudo code
\begin{align*}
  \OPH &=  \TRANSFER[\EXPR]\EXPR\EXPR\EXPR\EXPR \\
       & \MATCH\ {status\ \OPH}\ \WITH \\
       & \mid \STATUSPENDING \to \textrm{not yet included, continue waiting} \\
       & \mid \STATUSINCLUDING (n) \to \textrm{included with $n$
         blocks confirmation}\\
       & \mid \STATUSTIMEOUT \to \textrm{lost, try again}
\end{align*}
Grammar of types
\begin{align*}
  \TYPE & ::= \TADDR \mid \TOPH \mid \TPUH \mid \TPUK \mid \TCODE\ \TYPE \mid
          \TINT \mid \TUNIT \mid \TBOOL \mid \TSTRING \mid \TSTATUS \\
  & \mid \TYPE\to\TYPE \mid \TPAIR\ \TYPE\ \TYPE \mid \TLIST\ \TYPE
    \mid \TSUM\ \TYPE\ \TYPE \mid \TOPTION\ \TYPE \mid \TEXCEPTION
\end{align*}

\begin{definition}%[State of a node]
  The \emph{state of a node} is a tuple
  $\NODE = [ \EXPRS, \ACCOUNTS
  % , \OPERATIONS
  ]$, where $\EXPRS$ is a
  set of programs and $\ACCOUNTS \subseteq \TPAK \times \TPUK$  is a set of
  implicit accounts.
  % , and
  % $\OPERATIONS \subseteq \ECS \times \TOP $ is a set of pairs of
  % expression contexts and operations.
\end{definition}

\subsection{Global}
\label{sec:global}

\begin{definition}%[Manager]
A \emph{manager} manages a single implicit account. Managers are
represented by a partial mapping $\MANAGERS : \TPUK \partialto \TBAL
\times \TCOU$. If $\MANAGERS (\PUK) = \Angle{\BAL, \COU}$ is defined, then  $\PUK$ is the
public key of an account, $\BAL$ is its
balance and $\COU$ is its counter whose form is a value-flag pair
$(n,b) \in \Nat\times\TBOOL$, where $n$ is the value of the counter
and ${b}$ is its flag.   
\end{definition}

\begin{definition}%[Contractor]
  A \emph{contractor} manages a smart contract. Contractors are
  represented by a partial mapping $\CONTRACTORS : \TPUH \partialto
  \TCODE \times \TIME  \times \TBAL \times  \TSTORAGE$. If $\CONTRACTORS (\PUH) = \langle
  \CODE, \TIME, \BAL, \STORAGE \rangle $ is defined, then $\PUH$ is the
  public hash of a contract,
  $\CODE$ is its  code,
  $\TIME$ is the time when it was accepted,
  $\BAL$ is its current balance,
  and $\STORAGE$ is its current storage.
\end{definition}

When an operation is injected in a node, it enters a \emph{pending
  pool}. Later on, such a pending operation is either dismissed or
promoted to an accepted operation.

 %\begin{definition}%[Pending operation]
%An \emph{operation} is represented by a tuple  $\langle  \OP, \TIME, \STATUS \rangle $, where $\OP$ is an operation, $\TIME$ is the time when the operation was injected, and $\STATUS$ is either $\STATUSPENDING$,$\STATUSINCLUDED\ \TIME'$, or $\STATUSTIMEOUT$.  
%\end{definition}
\begin{definition}
  The pool of operations, $\PENDING$, is defined as a partial mapping from
  operation hashes $\OPH$: 
  \begin{align*}
    \PENDING &= \TOPH \partialto \TOP\times \TTIME \times \TSTATUS \\
    \TSTATUS&= \STATUSPENDING + \STATUSINCLUDED\ \TTIME + \STATUSTIMEOUT
  \end{align*}

  if $\PENDING(\OPH) = \langle  \OP, \TIME, \STATUS
\rangle $ is defined, then $\OPH$ is the public hash of the operation, $\OP$ is the operation, $\TIME$ is the time when the
operation was injected, and $\STATUS$ is either $\STATUSPENDING$,
$\STATUSINCLUDED\ \TIME'$, or $\STATUSTIMEOUT$.
A pool $\PENDING$ is \emph{well-formed} if, for all $\OPH$, $\PENDING (\OPH) =
  \Angle{\OP,\TIME, \STATUSINCLUDED\ \TIME'}$ implies $\TIME' \ge
  \TIME$ and $\OPH = \GENERATEOPH (\OP, \TIME)$.
\end{definition}

A \emph{pending operation} is represented by
$\OPH \mapsto \langle  \OP, \TIME, \STATUSPENDING\rangle $,
where $\OP$ is an operation, $\OPH$ is the operation hash,
and $\TIME$ is the time when the operation was injected.
Once it is accepted, it changes to an \emph{accepted operation}:
$\OPH \mapsto \langle  \OP, \TIME, \STATUSINCLUDED\ \TIME'\rangle $, where
$\TIME' \ge \TIME$ is  when the operation was included in the
blockchain.  The operation can also be dropped at any time, which is
represented by
$\OPH \mapsto \langle  \OP, \TIME, \STATUSTIMEOUT\rangle $. There are several
causes for dropping, primarily timeout or overflow of the pending
pool which is limited in size in the implementation.

% After some time, a pending operation may be included in the blockchain as an accepted operation.

 % \begin{definition}%[Accepted operation]
% An \emph{accepted operation}  is a tuple $\langle  \OP, \OPH, \TIME, \TIME' \rangle
% $, where $\OP$ is an operation, $\OPH$ is the operation hash, $\TIME$ is the time when the operation was injected, and
% $\TIME'$ is the time when it was included in the blockchain. 
% \end{definition}

\begin{definition}%[Blockchain]
  The \emph{state of a blockchain} is a tuple
  $\BLOCKCHAIN = [\PENDING, \MANAGERS, \CONTRACTORS, \TIME]$ where
  $\PENDING$ is a pool of operations, $\MANAGERS$ is a map of managers,
  $\CONTRACTORS$ is a map of contractors, and $\TIME$ is the current
  time. 
\end{definition}
\begin{definition}%[Blockchain system]
A \emph{blockchain configuration} is a pair
$ \NODE \| \BLOCKCHAIN$ where
\fxnote[inline]{Should the model contain multiple nodes?}
\begin{enumerate}
\item $\NODE = [\EXPRS, \ACCOUNTS]$ is the state of a node, and
\item $\BLOCKCHAIN = [\PENDING, \MANAGERS, \CONTRACTORS,
  \TIME]$ is the state of a blockchain  
  such that
  \begin{itemize}
  \item
    $\forall a \in \ACCOUNTS \implies a.\PUK \in \DOM( \MANAGERS)$;
  \end{itemize}
\end{enumerate}
\end{definition}
\begin{definition}
  Preserved by a step on a well-formed configuration $ [\EXPRS, \ACCOUNTS] \|
  [\PENDING, \MANAGERS, \CONTRACTORS, \TIME]$:
  \todo[inline]{Is this preserved?}
  \begin{itemize}
  \item for all $\OPH$ in $\EXPRS$, $\OPH \in \DOM (\PENDING)$,
  \item for all $\PUK$ in $\EXPRS$, $\PUK \in \DOM (\MANAGERS)$,
  \item for all $\PUH$ in $\EXPRS$, $\PUH \in \DOM (\CONTRACTORS)$.
  \end{itemize}
\end{definition}

\clearpage{}
\subsection{Properties of blockchain state transitions}

If $[\PENDING, \MANAGERS, \CONTRACTORS, \TIME] \BlockTrans{}
[\PENDING', \MANAGERS', \CONTRACTORS', \TIME']$, then
\begin{itemize}
\item $\TIME \le \TIME'$
\item $\DOM (\PENDING) \subseteq \DOM(\PENDING')$
\item for all $\OPH \in \DOM(P)$, if 
  $\PENDING (\OPH) = \Angle{\OP,  \hat\TIME, \STATUS}$, then either
  \begin{itemize}
  \item     $\PENDING'    (\OPH) = \PENDING (\OPH)$; or
  \item $\STATUS  = \STATUSPENDING$ and $\PENDING' (\OPH) =
    \Angle{\OP, \hat\TIME, \STATUSTIMEOUT}$; or
  \item  $\STATUS  = \STATUSPENDING$, $\TIME - \hat\TIME
    \le 60 $, and
    $\PENDING' (\OPH) =       \Angle{\OP, \hat\TIME, \STATUSINCLUDED\
      \TIME}$.
  \end{itemize}
 \item for all $\OPH \in \DOM(P)$ and $\PENDING (\OPH) = \Angle{\OP,  \hat\TIME, \STATUSPENDING}$, 
    \begin{itemize}
    	\item  if $\MANAGERS(\OP.\PUK) =  \Angle{\BAL, \COU}$ then  
  $\COU.b = \TRUE$.
    \end{itemize}
 \item $ \DOM (\MANAGERS) \subseteq \DOM (\MANAGERS')$
 \item for all $\PUK \in \DOM (\MANAGERS)$\\
   if $\MANAGERS (\PUK) =
   \Angle{\BAL, \COU}$, 
   then $\MANAGERS' (\PUK) =
   \Angle{\BAL', \COU'}$ and
   \begin{itemize}
   \item if $\COU.b=\TRUE$ and $\COU'.b=\FALSE$, then $\COU.n' \in \{
     \COU.n,  \COU.n+1\}$,
   \item otherwise $\COU.n = \COU'.n$
   \item If $\COU.n = \COU'.n$, then $\BAL = \BAL'$.
   % \item if $\COU.b= \FALSE$, then $n' = n$
   % \item if $\COU.b= \TRUE$ and $b'=\TRUE$, then $n' =n$
   \end{itemize}

 \item $ \DOM (\CONTRACTORS) \subseteq \DOM (\CONTRACTORS')$
   \begin{itemize}
   \item for all $\PUH \in \DOM (\CONTRACTORS)$,
     $\CONTRACTORS (\PUH).\CODE = \CONTRACTORS' (\PUH).\CODE$
   \end{itemize}
\end{itemize}

\subsection{Typing related properties}

To describe contracts we associate each public hash to a pair of
types, the parameter type and the storage type.
\begin{align*}
  \Delta & ::= \EmptyEnv \mid \PUH: \TPAIR\ \TYPE\ \TYPE, \Delta
\end{align*}

Define $ \Delta \vdash \CONTRACTORS$ by
\begin{itemize}
\item $\DOM (\Delta) = \DOM (\CONTRACTORS)$
\item $\CONTRACTORS (\PUH).\CODE : \Delta (\PUH) \to FIXEDOUTPUTTYPE$
\item $\CONTRACTORS (\PUH).\STORAGE : \TYPE_s$ where $\Delta (\PUH) =
  \TPAIR\ \TYPE_p\ \TYPE_s$. 
\end{itemize}

Preservation

If $[\PENDING, \MANAGERS, \CONTRACTORS, \TIME] \BlockTrans{}
[\PENDING', \MANAGERS', \CONTRACTORS', \TIME']$
and $\Delta \vdash \CONTRACTORS$, \\
then
there is some $\Delta' \supseteq \Delta$ such that $\Delta' \vdash
\CONTRACTORS'$.

Consider a blockchain state  $[\PENDING, \MANAGERS, \CONTRACTORS,
\TIME]$ where $\Delta \vdash \CONTRACTORS$.

For all $\OPH\in\DOM (\PENDING)$,
if  $\PENDING (\OPH) =  \Angle{\OP, \hat\TIME, \STATUS}$
\begin{itemize}
\item if $\OP = \TRANSFER\NTEZ\PUK{\PUK'}\MTEZ$, then
\begin{itemize}
  \item
  $\PUK, \PUK' \in \DOM (\MANAGERS)$
\end{itemize}
\item if $\OP = \TRANSFER[\PARAMETER]\NTEZ\PUK\PUH\MTEZ$, then
  \begin{itemize}
  \item $\PUK \in \DOM (\MANAGERS)$ and $\PUH\in\DOM (\CONTRACTORS)$
  \item $\PARAMETER : \TYPE_p$ where $\Delta (\PUH)  =\TPAIR\ \TYPE_p\ \TYPE_s$. 
  \end{itemize}

\item if $\OP = \ORIGINATE\NTEZ\PUK\CODE\STRING\MTEZ$ and $\STATUS =
  \STATUSINCLUDED\ \TIME'$, then
   \begin{itemize}
   \item $\PUK \in \DOM (\MANAGERS)$
  \item $\CODE : \TPAIR\ \TYPE_p\ \TYPE_s$
  \item $\STRING : \TYPE_s$
  \item if $\STATUS =
  \STATUSINCLUDED\ \TIME'$, then $\PUH =\GENERATEHASH (\CODE, \hat\TIME) \in \DOM (\CONTRACTORS)$ and $\Delta (\PUH)  =\TPAIR\ \TYPE_p\ \TYPE_s$ (or $\CONTRACTORS(\PUH).\CODE = \CODE$).

  \end{itemize}

 \item if $st = \STATUSPENDING$ and $\MANAGERS(\PUK) = \Angle{\BAL,
     \COU}$, then $\BAL \ge \NTEZ +  \MTEZ$ 
\end{itemize}

\section{Transitions}
\label{sec:transitions}

We distinguish several kinds of transitions depending on their scope:
\begin{enumerate}
\item $\ExprTrans$ single-step evaluation of expressions in the node state,
\item $\NodeTrans$ internal transitions of a node,
\item $\BlockTrans$ transitions of the blockchain state,
\item $\SystemTrans$ blockchain system transitions.
\end{enumerate}

Evaluation of expressions is standard call-by-value lambda calculus defined
using evaluation contexts $\EC{}$.
\begin{mathpar}
  \inferrule{}{
    \EC{(\lambda x.\EXPR)\VAL} \ExprTrans \EC{\EXPR[\VAL/x]}
  }

  \inferrule{}{
    \EC{\FST (\VAL_1, \VAL_2)} \ExprTrans \EC{\VAL_1}
  }

  \inferrule{}{
    \EC{\SND (\VAL_1, \VAL_2)} \ExprTrans \EC{\VAL_2}
  }

  \inferrule{
    \TRY \notin \EC[F]{}
  }{
    \EC{\TRY\ {\EC[F]{\RAISE\ \VAL}}\ \EXCEPT\ \EXPR} \ExprTrans
    \EC{\EXPR\ \VAL}
  }

  \inferrule{}{
    \EC{\MATCH\ \NIL\ \WITH\ \NIL\to\EXPR \mid \dots}
    \ExprTrans
    \EC{\EXPR}
  }

  \inferrule{}{
    \EC{\MATCH\ (\CONS\ \VAL_1\ \VAL_2)\ \WITH\ (\CONS\ x_1\ x_2)\to\EXPR \mid \dots}
    \ExprTrans
    \EC{\EXPR[\VAL_1,\VAL_2/x_1, x_2]}
  }

  \inferrule{}{
    \EC{\MATCH\ \NONE\ \WITH\ \NONE\to\EXPR \mid \dots}
    \ExprTrans
    \EC{\EXPR}
  }

  \inferrule{}{
    \EC{\MATCH\ (\SOME\ \VAL)\ \WITH\ (\SOME\ x)\to\EXPR \mid \dots}
    \ExprTrans
    \EC{\EXPR[\VAL/x]}
  }

  \inferrule{}{
    \EC{\MATCH\ (\LEFT\ \VAL)\ \WITH\ (\LEFT\ x)\to\EXPR \mid \dots}
    \ExprTrans
    \EC{\EXPR[\VAL/x]}
  }

  \inferrule{}{
    \EC{\MATCH\ (\RIGHT\ \VAL)\ \WITH\ (\RIGHT\ x)\to\EXPR \mid \dots}
    \ExprTrans
    \EC{\EXPR[\VAL/x]}
  }
\end{mathpar}

The internal transitions of a node are just evaluation of expressions.
\begin{mathpar}
  \inferrule[Node-Eval]
  {
    \EXPR \ExprTrans \EXPR'
  }{
    [\EC\EXPR :: \EXPRS, \ACCOUNTS] \NodeTrans{}
    [\EC{\EXPR'} :: \EXPRS, \ACCOUNTS]
  }
\end{mathpar}

\begin{mathpar}
  \inferrule[Node-Inject]{
    \Angle{\PAK,\PUK} \in \ACCOUNTS \\
    \CHECKBAL (\MANAGERS, \PUK, \NTEZ, \MTEZ) \\
    \CHECKCOU (\MANAGERS, \PUK) \\
    \CHECKPUB (\MANAGERS, \PUK') \\
    \CHECKGAS (\PUK, \PUK', \NTEZ, \MTEZ) \\
    %\OPH = \GENERATEOPH (\PUK, \PUK','', \NTEZ, \MTEZ, \TIME) \\
    \OPH = \GENERATEOPH (\OP, \TIME) \\
    \OP = \TRANSFER\NTEZ\PUK{\PUK'}\MTEZ    
  }{
    { [\EC\OP :: \EXPRS, \ACCOUNTS] \|
      [\PENDING, \MANAGERS, \CONTRACTORS, \TIME] } \SystemTrans \\
   { [\EC{\OPH}  :: \EXPRS, \ACCOUNTS] \|
     [ \OPH \mapsto \Angle{\OP, \TIME, \STATUSPENDING}
     ::\PENDING,
     \UPDATECOU(\MANAGERS, \PUK, \TRUE),
     \CONTRACTORS,
     \TIME]
   }
  }
\end{mathpar}

\begin{mathpar}
  \inferrule[Block-Accept]{
    \OP = \TRANSFER\NTEZ\PUK{\PUK'}\MTEZ \\
    \TIME - \hat\TIME \le 60
  }{
    { 
      [\OPH \mapsto \Angle{\OP, \hat \TIME, \STATUSPENDING}
     ::\PENDING, \MANAGERS,
      \CONTRACTORS, \TIME]}
    \BlockTrans  \\
    { 
      [\OPH \mapsto \Angle{\OP, \hat\TIME, \STATUSINCLUDING\ \TIME} :: \PENDING, 
      \UPDATESUCC(\MANAGERS, \PUK, \PUK', \NTEZ, \MTEZ),
      \CONTRACTORS, \TIME +1]}
  }
\end{mathpar}

The timeout transition applies to all operations.
\begin{mathpar}
  \inferrule[Block-Timeout]{
    % \OP = \TRANSFER[\STRING]\NTEZ\PUK\ADDR\MTEZ \\
    \TIME-\hat\TIME > 60
  }{ 
    {[\OPH \mapsto \Angle{\OP, \hat \TIME, \STATUSPENDING}
     ::\PENDING, \MANAGERS,
      \CONTRACTORS, \TIME]}
    \BlockTrans \\
    { 
      [\OPH \mapsto \Angle{\OP, \hat \TIME, \STATUSTIMEOUT}
     :: \PENDING, \UPDATECOU(\MANAGERS, \OP.\PUK, \FALSE),
      \CONTRACTORS, \TIME]}
  }
\end{mathpar}

% \begin{itemize}
% \item library call \texttt{call\_contract} with \texttt{confirmations:None} corresponds to
%   \begin{enumerate}
%   \item{} first \TirName{Node-Op'}, \TirName{Node-Inject'}
%   \item{} first \TirName{Node-Op'}, \TirName{Node-Reject(balance,count,pubkey,fee)}
%   \end{enumerate}
% \item afterwards must check the status to find whether the
%   transfer is still pending or \TirName{Block-Accept''} or
%   \TirName{Block-Timeout''} has happened.
% \item no further transfer is possible before the status is resolved to
%   accept or timeout.
% \end{itemize}

\subsection{Transitions on Nodes}
\label{sec:transitions-nodes}

Each node has (nondeterministic) rules to propose an operation.
When an operation $\OP$ appears, we check that the account is local by
looking up its public key $\OP.\PUK$ in the local accounts $\ACCOUNTS$ and
consider it signed with the corresponding private key $\PAK$.

\begin{mathpar}
  \inferrule[Node-System]
  {\NODE \NodeTrans \NODE'}
  { {\NODE :: \overline\NODE \| \BLOCKCHAIN}
    \SystemTrans
    {\NODE' :: \overline\NODE \| \BLOCKCHAIN}}

  \inferrule[Block-System]
  {\BLOCKCHAIN \BlockTrans \BLOCKCHAIN'}
  { {\overline\NODE \| \BLOCKCHAIN}
    \SystemTrans
    {\overline\NODE\| \BLOCKCHAIN'}}
\end{mathpar}

\begin{mathpar}
  \inferrule[rejected of counter]{
    \NEG\ \CHECKCOU (\BLOCKCHAIN.\MANAGERS, \OP.\PUK) \\
    }{
    { [\EC\OP  :: \EXPRS, \ACCOUNTS] \|
      \BLOCKCHAIN
    }
    \SystemTrans
    { [\EC{\RAISE\ \ERRCOUNT} :: \EXPRS, \ACCOUNTS] \|
      \BLOCKCHAIN
   }
  }
\end{mathpar}

\begin{mathpar}
  \inferrule[rejected of balance]{
    \NEG\ \CHECKBAL (\BLOCKCHAIN.\MANAGERS, \OP.\PUK, \OP.\NTEZ, \OP.\MTEZ) \\
    }{
    { [\EC\OP :: \EXPRS, \ACCOUNTS] \|
      \BLOCKCHAIN
      } \SystemTrans
    { [\EC{\RAISE\ \ERRBAL} :: \EXPRS, \ACCOUNTS] \|
      \BLOCKCHAIN
    }
  }
\end{mathpar}

\subsection{An Implicit Account Transfer}
\iffalse 
\fi

\begin{mathpar}
  \inferrule[rejected of public key]{
    \NEG\ \CHECKPUB (\BLOCKCHAIN.\MANAGERS, \PUK') \\
    \OP = \TRANSFER\NTEZ\PUK{\PUK'}\MTEZ \\
  }{
    { [\EC\OP :: \EXPRS, \ACCOUNTS] \|
      \BLOCKCHAIN
    } \SystemTrans
   { [\EC{\RAISE\ \ERRPUK} :: \EXPRS, \ACCOUNTS] \|
     \BLOCKCHAIN
    }
  }
\end{mathpar}

\begin{mathpar}
  \inferrule[rejected of fee]{
    \NEG\ \CHECKGAS (\BLOCKCHAIN.\MANAGERS, \PUK,\PUK',\NTEZ, \MTEZ) \\
    \OP = \TRANSFER\NTEZ\PUK{\PUK'}\MTEZ \\
  }{
    { [\EC\OP :: \EXPRS, \ACCOUNTS] \|
      \BLOCKCHAIN
     } \SystemTrans
    { [\EC{\RAISE\ \ERRFEE} :: \EXPRS, \ACCOUNTS] \|
      \BLOCKCHAIN
    }
  }
\end{mathpar}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\iffalse 
Rule 6 [included]:
\begin{mathpar}
  \inferrule[Block-Accept]{
    \OP = \TRANSFER\NTEZ\PUK{\PUK'}\MTEZ \\
    \TIME'-\TIME < 60
  }{
    {[\Angle{\OP,\OPH, \TIME} :: \PENDING, \ACCEPTED, \MANAGERS,
      \CONTRACTORS, \TIME']}
    \BlockTrans  \\
    {[\PENDING, \Angle{\OP,\OPH, \TIME,\TIME'} :: \ACCEPTED,
      \UPDATESUCC(\MANAGERS, \PUK, \PUK', \NTEZ, \MTEZ),
      \CONTRACTORS, \TIME'+1]}
  }
\end{mathpar}
Rule 7 [timeout]: (applies to both, implicit transfers and contract invocations)
\begin{mathpar}
  \inferrule[Block-Timeout]{
    \OP = \TRANSFER[\STRING]\NTEZ\PUK\ADDR\MTEZ \\
    \TIME'-\TIME \ge 60
  }{
    {[\Angle{\OP,\OPH, \TIME} :: \PENDING, \ACCEPTED, \MANAGERS,
      \CONTRACTORS, \TIME']}
    \BlockTrans
    {[\PENDING, \ACCEPTED, \UPDATECOU(\MANAGERS, \OP.\PUK, \FALSE),
      \CONTRACTORS, \TIME']}
  }
\end{mathpar}
\fi
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Smart Contracts}
A. Originate

\begin{mathpar}
  \inferrule[Block-Originate]{
   \Angle{\PAK,\PUK} \in \ACCOUNTS \\ \CHECKBAL (\MANAGERS, \PUK, \NTEZ, \MTEZ) \\
    \CHECKCOU (\MANAGERS, \PUK) \\
    \CHECKPRG (\CODE) \\
    \CHECKGAS (\CODE, \INIT, \NTEZ, \MTEZ)  \\
    \CHECKINIT (\CODE, \STRING) \\
    %\OPH = \GENERATEOPH(\PUK, \CODE, \STRING, \NTEZ, \MTEZ, \TIME) \\
    \OPH = \GENERATEOPH(\OP, \TIME) \\
    \OP = \ORIGINATE\NTEZ\PUK\CODE\STRING\MTEZ    
  }{
    [\EC\OP  :: \EXPRS, \ACCOUNTS,
      ] \| [\PENDING, \MANAGERS, \CONTRACTORS, \TIME] \to \\
    [\EC{\OPH}  :: \EXPRS, \ACCOUNTS] \|
      [\OPH \mapsto \Angle{\OP, \TIME, \STATUSPENDING}
     ::\PENDING, \UPDATECOU(\MANAGERS,\PUK,\TRUE), \CONTRACTORS, \TIME]
  }
\end{mathpar}

\begin{mathpar}
  \inferrule[rejected of code]{
    \NEG\ \CHECKPRG (\CODE) \\
    \OP = \ORIGINATE\NTEZ\PUK\CODE\STRING\MTEZ
  }{
    [\EC\OP :: \EXPRS, \ACCOUNTS] \| \BLOCKCHAIN \to 
    [\EXPRS[\RAISE\ \ERRPRG] :: \EXPRS, \ACCOUNTS] \|
     \BLOCKCHAIN
  }
\end{mathpar}

\iffalse 
Rule 3 [rejected of counter]:
\begin{mathpar}
  \inferrule{
    \NEG\ \CHECKCOU (\BLOCKCHAIN.\MANAGERS, \PUK) \\
    \OP = \ORIGINATE\NTEZ\PUK\CODE\STRING\MTEZ
  }{
    [\EC\OP :: \EXPRS, \ACCOUNTS] \| \BLOCKCHAIN \to 
    [\EXPRS[\RAISE\ \ERRCOUNT] :: \EXPRS, \ACCOUNTS] \|
     \BLOCKCHAIN
  }
\end{mathpar}

\begin{mathpar}
  \inferrule[rejected of balance]{
    \NEG\ \CHECKBAL (\BLOCKCHAIN.\MANAGERS, \PUK, \NTEZ, \MTEZ) \\
    \OP = \ORIGINATE\NTEZ\PUK\CODE\STRING\MTEZ
  }{
    [\EC\OP :: \EXPRS, \ACCOUNTS] \| \BLOCKCHAIN \to 
    [\EXPRS[\RAISE\ \ERRBAL] :: \EXPRS, \ACCOUNTS] \|
     \BLOCKCHAIN
  }
\end{mathpar}
\fi

\begin{mathpar}
  \inferrule[rejected of fee]{
    \NEG\ \CHECKGAS (\CODE, \INIT, \NTEZ, \MTEZ) \\
    \OP = \ORIGINATE\NTEZ\PUK\CODE\STRING\MTEZ
  }{
    [\EC\OP :: \EXPRS, \ACCOUNTS] \| \BLOCKCHAIN \to 
    [\EXPRS[\RAISE\ \ERRFEE] :: \EXPRS, \ACCOUNTS] \|
     \BLOCKCHAIN
  }
\end{mathpar}

\begin{mathpar}
  \inferrule[rejected of initial storage]{
    \NEG\ \CHECKINIT (\CODE, \INIT) \\
    \OP = \ORIGINATE\NTEZ\PUK\CODE\STRING\MTEZ
  }{
    [\EC\OP :: \EXPRS, \ACCOUNTS] \| \BLOCKCHAIN \to 
    [\EXPRS[\RAISE\ \ERRINIT] :: \EXPRS, \ACCOUNTS] \|
     \BLOCKCHAIN
  }
\end{mathpar}

\begin{mathpar}
  \inferrule[Block-Accept]{
    \OP = \ORIGINATE\NTEZ\PUK\CODE\STRING\MTEZ \\
    \PUH = \GENERATEHASH(\CODE, \TIME) \\
    \TIME-\hat\TIME  \le 60
  }{
      [\EC{\OPH}  :: \EXPRS, \ACCOUNTS] \| [\OPH \mapsto \Angle{\OP, \hat\TIME, \STATUSPENDING} :: \PENDING, \MANAGERS, \CONTRACTORS, \TIME]
    \SystemTrans \\
     [\EC{ \PUH}  :: \EXPRS, \ACCOUNTS] \|  [\OPH \mapsto \Angle{\OP, \hat\TIME, \STATUSINCLUDING\ \TIME} :: \PENDING, \UPDATESUCC
      (\MANAGERS, \PUK, \NTEZ, \MTEZ),\\ \PUH \mapsto  \Angle{\CODE, \TIME, \NTEZ, \STRING} :: \CONTRACTORS, \TIME+1]
  }
\end{mathpar}

% \begin{mathpar}
%   \inferrule[Block-Timeout]{
%     %\OP = \ORIGINATE\NTEZ\PUK\CODE\STRING\MTEZ \\
%     \TIME-\hat\TIME > 60
% }{
%       [ \OPH \mapsto \Angle{\OP, \hat\TIME, \STATUSPENDING}
%      ::\PENDING, \MANAGERS,
%       \CONTRACTORS, \TIME]  
%       \BlockTrans \\    
%       [ \OPH \mapsto \Angle{\OP, \hat\TIME, \STATUSTIMEOUT}
%      :: \PENDING, \UPDATECOU(\MANAGERS, \OP.\PUK, \FALSE),
%       \CONTRACTORS, \TIME] 
%   }

% \end{mathpar}
~\\
~\\
B. \textbf{Transfer}

\begin{mathpar}
  \inferrule[Block-Call]{
    \CHECKBAL (\MANAGERS, \PUK, \NTEZ, \MTEZ) \\ 
    \CHECKCOU (\MANAGERS, \PUK) \\
    \CHECKPUH (\CONTRACTORS, \PUH) \\
    \CHECKARG (\CONTRACTORS, \PUH, \PARAMETER) \\
    \CHECKGAS (\CONTRACTORS, \PUH, \PARAMETER, \MTEZ) \\
%    \OPH = \GENERATEOPH (\PUK,\PUH, \PARAMETER,\NTEZ, \MTEZ, \TIME) \\
    \OPH = \GENERATEOPH (\OP, \TIME) \\
    \OP =  \TRANSFER[\PARAMETER]\NTEZ\PUK{\PUH}\MTEZ \\    
  }{ [\EC\OP  :: \EXPRS, \ACCOUNTS,
      ] \| [\PENDING, \MANAGERS, \CONTRACTORS, \TIME]     
    \SystemTrans \\
    [\EC{\OPH}  :: \EXPRS, \ACCOUNTS] \|
      [\OPH \mapsto \Angle{\OP, \TIME, \STATUSPENDING}
     :: \PENDING, \UPDATECOU(\MANAGERS, \PUK, \TRUE), \CONTRACTORS, \TIME]}
\end{mathpar}

\iffalse 
Rule 2 [rejected of counter]:
\begin{mathpar}
  \inferrule{
    \NEG\ \CHECKCOU (\BLOCKCHAIN.\MANAGERS, \PUK) \\
    \OP = \TRANSFER[\PARAMETER]\NTEZ\PUK{\PUH}\MTEZ \\
  }{[\EC\OP :: \EXPRS, \ACCOUNTS] \| \BLOCKCHAIN      
    \SystemTrans
   [\EC[\RAISE\ \ERRCOUNT] :: \EXPRS, \ACCOUNTS] \| \BLOCKCHAIN }
\end{mathpar}
~\\
~\\
Rule 3 [rejected of balance]:
\begin{mathpar}
  \inferrule{
    \NEG\ \CHECKBAL (\MANAGERS, \PUK, \NTEZ, \MTEZ) \\
    \OP = \TRANSFER[\PARAMETER]\NTEZ\PUK{\PUH}\MTEZ \\
  }{[\EXPRS, \ACCOUNTS, (\EC{},\OP) :: \OPERATIONS] \|
      [\PENDING, \ACCEPTED, \MANAGERS, \CONTRACTORS, \TIME]      
    \SystemTrans
    [\EXPRS, \ACCOUNTS, \OPERATIONS] \|
      [\PENDING, \ACCEPTED, \MANAGERS, \CONTRACTORS, \TIME]}
\end{mathpar}
\fi

\begin{mathpar}
  \inferrule[rejected of public hash]{
    \NEG\ \CHECKPUH (\BLOCKCHAIN.\CONTRACTORS, \PUH) \\
    \OP = \TRANSFER[\PARAMETER]\NTEZ\PUK{\PUH}\MTEZ
  }{[\EC\OP  :: \EXPRS, \ACCOUNTS] \| \BLOCKCHAIN    
    \SystemTrans
    [\EXPRS[\RAISE\ \ERRPUH], \ACCOUNTS] \| \BLOCKCHAIN}
\end{mathpar}

\begin{mathpar}
  \inferrule[rejected of argument]{
    \NEG\ \CHECKARG (\BLOCKCHAIN.\CONTRACTORS, \PUH, \PARAMETER)\\
    \OP = \TRANSFER[\PARAMETER]\NTEZ\PUK{\PUH}\MTEZ
  }{[\EC\OP  :: \EXPRS, \ACCOUNTS] \| \BLOCKCHAIN    
    \SystemTrans
    [\EXPRS[\RAISE\ \ERRARG], \ACCOUNTS] \| \BLOCKCHAIN}
\end{mathpar}

\begin{mathpar}
  \inferrule[rejected of fee]{
    \NEG\ \CHECKGAS (\BLOCKCHAIN.\CONTRACTORS, \PUH, \PARAMETER, \MTEZ) \\ \OP = \TRANSFER[\PARAMETER]\NTEZ\PUK{\PUH}\MTEZ 
  }{[\EC\OP  :: \EXPRS, \ACCOUNTS] \| \BLOCKCHAIN    
    \SystemTrans
    [\EXPRS[\RAISE\ \ERRFEE], \ACCOUNTS] \| \BLOCKCHAIN}
\end{mathpar}

\begin{mathpar}
  \inferrule[block Accept]{
    \OP = \TRANSFER[\PARAMETER]\NTEZ\PUK\PUH\MTEZ \\
    \TIME-\hat\TIME \le 60
  }{
    [\OPH \mapsto \Angle{\OP, \hat\TIME, \STATUSPENDING} :: \PENDING, \MANAGERS,
    \CONTRACTORS, \TIME]
    \BlockTrans \\
    [\OPH \mapsto \Angle{\OP, \hat\TIME, \STATUSINCLUDING\ \TIME} :: \PENDING, 
    \UPDATESUCC (\MANAGERS, \PUK, '', \NTEZ, \MTEZ), \\
    \UPDATECONSTR (\CONTRACTORS, \PUH, \NTEZ, \PARAMETER),
    \TIME+1]
  }
\end{mathpar}
\section{Queries}
\begin{mathpar}
  \inferrule[counter]{
    \BLOCKSYSTEM = 
      [\PENDING, \PUK \mapsto \Angle{\BAL, \COU} :: \MANAGERS, \CONTRACTORS, \TIME]
  }{[\EC{\GETCOUNTER\PUK} :: \EXPRS, \ACCOUNTS]  \| \BLOCKCHAIN
    \NodeTrans\ [\EC{\COU} ::\EXPRS, \ACCOUNTS]  \| \BLOCKCHAIN}
\end{mathpar}

\begin{mathpar}
  \inferrule[balance of an implicit account]{
    \BLOCKCHAIN = 
      [\PENDING, \PUK \mapsto \Angle{\BAL, \COU} :: \MANAGERS, \CONTRACTORS, \TIME]
  }{[\EC{\GETBALANCE\PUK} :: \EXPRS, \ACCOUNTS]  \| \BLOCKCHAIN
    \NodeTrans\ [\EC{\BAL} ::\EXPRS, \ACCOUNTS]  \| \BLOCKCHAIN}
\end{mathpar}

\begin{mathpar}
\inferrule[balance of smart contract]{
    \BLOCKSYSTEM = 
      [\PENDING,  \MANAGERS, \PUH \mapsto \Angle{\CODE, \TIME, \BAL, \STORAGE} :: \CONTRACTORS, \TIME]
  }{[\EC{\GETBALANCE\PUH} :: \EXPRS, \ACCOUNTS]  \| \BLOCKCHAIN
    \NodeTrans\ [\EC{\BAL} ::\EXPRS, \ACCOUNTS]  \| \BLOCKCHAIN}
\end{mathpar}

\begin{mathpar}
  \inferrule[storage]{
    \BLOCKSYSTEM = 
      [\PENDING, \MANAGERS, \PUH \mapsto \Angle{\CODE, \TIME, \BAL, \STORAGE} :: \CONTRACTORS, \TIME]
  }{[\EC{\GETSTORAGE\PUH} :: \EXPRS, \ACCOUNTS]  \| \BLOCKCHAIN
    \NodeTrans\ [\EC{\STORAGE} ::\EXPRS, \ACCOUNTS]  \| \BLOCKCHAIN}
\end{mathpar}

\begin{mathpar}
  \inferrule[code]{
    \BLOCKSYSTEM = 
      [\PENDING,  \MANAGERS, \PUH \mapsto \Angle{\CODE, \TIME, \BAL, \STORAGE} :: \CONTRACTORS, \TIME]
  }{[\EC{\GETCODE\PUH} :: \EXPRS, \ACCOUNTS]  \| \BLOCKCHAIN
    \NodeTrans\ [\EC{\CODE} ::\EXPRS, \ACCOUNTS]  \| \BLOCKCHAIN}
\end{mathpar}

\begin{mathpar}
  \inferrule[type]{
    \BLOCKSYSTEM = 
      [\PENDING,  \MANAGERS, \PUH \mapsto \Angle{\CODE, \TIME, \BAL, \STORAGE} :: \CONTRACTORS, \TIME]
  }{[\EC{\GETTYPE\PUH} :: \EXPRS, \ACCOUNTS]  \| \BLOCKCHAIN
    \NodeTrans\ [\EC{\Delta (\PUH)} ::\EXPRS, \ACCOUNTS]  \| \BLOCKCHAIN}
\end{mathpar}

\begin{mathpar}
  \inferrule[fail of a query for an implicit account]{
    \NEG\ \CHECKPUB (\BLOCKCHAIN.\MANAGERS, \PUK) \
  }{[\EC{\GETBALANCE\PUK/\GETCOUNTER\PUK} :: \EXPRS, \ACCOUNTS]  \| 
  \BLOCKCHAIN 
    \NodeTrans {[\EC{\RAISE\ \ERRPUK} ::\EXPRS, \ACCOUNTS]  \| \BLOCKCHAIN}}
\end{mathpar}

\begin{mathpar}
  \inferrule[fail of a query for a smart contract]{
    \NEG\ \CHECKPUH (\BLOCKCHAIN.\CONTRACTORS, \PUH) 
  }{[\EC{\GETBALANCE\PUH/\GETSTORAGE\PUH/\GETCODE\PUH/\GETTYPE\PUH} :: \EXPRS, \ACCOUNTS]  \| \BLOCKCHAIN
    \NodeTrans\\ [\EC{\RAISE\ \ERRPUH} ::\EXPRS, \ACCOUNTS]  \| \BLOCKCHAIN}
\end{mathpar}

\begin{mathpar}
  \inferrule[status-pending]{
    \BLOCKSYSTEM = 
      [\OPH \mapsto \Angle{\OP, \TIME, \STATUSPENDING} :: \PENDING,  \CONTRACTORS, \TIME]
  }{[\EC{\GETSTATUS\OPH} :: \EXPRS, \ACCOUNTS]  \| \BLOCKCHAIN
    \NodeTrans\ [\EC{\STATUSPENDING} ::\EXPRS, \ACCOUNTS]  \| \BLOCKCHAIN}
\end{mathpar}

\begin{mathpar}
  \inferrule[status-including]{
    \BLOCKSYSTEM = 
      [\OPH \mapsto \Angle{\OP, \hat\TIME, \STATUSINCLUDING\ \TIME'} :: \PENDING,  \CONTRACTORS, \TIME] \\
    $n$ \ = \TIME\ - \TIME'
  }{[\EC{\GETSTATUS\OPH} :: \EXPRS, \ACCOUNTS]  \| \BLOCKCHAIN
    \NodeTrans\ [\EC{\STATUSINCLUDING($n$)} ::\EXPRS, \ACCOUNTS]  \| \BLOCKCHAIN}
\end{mathpar}

\begin{mathpar}
  \inferrule[status-timeout]{ 
   \BLOCKSYSTEM = 
      [\OPH \mapsto \Angle{\OP, \hat\TIME, \STATUSTIMEOUT} :: \PENDING, \CONTRACTORS, \TIME]
  }{[\EC{\GETSTATUS\OPH} :: \EXPRS, \ACCOUNTS]  \| \BLOCKCHAIN
    \NodeTrans\ [\EC{\STATUSTIMEOUT} ::\EXPRS, \ACCOUNTS]  \| \BLOCKCHAIN}
\end{mathpar}
\section{Functions}
\begin{enumerate}
    \item Function $\CHECKPUB(\MANAGERS, \PUK)$ checks whether a public key $\PUK$ exists in $\MANAGERS$.
    \begin{itemize}
    \item[] $\CHECKPUB(\MANAGERS, \PUK)$ = $\TRUE$ iff $\PUK \in \DOM(\MANAGERS)$
    \end{itemize}
    \item Function $\CHECKPUH(\CONTRACTORS, \PUH)$ checks whether a public hash $\PUH$ exists in $\CONTRACTORS$.
     \begin{itemize}
    \item[] $\CHECKPUH(\CONTRACTORS, \PUH)$ = $\TRUE$ iff $\PUH \in \DOM(\CONTRACTORS)$
    \end{itemize}
    \item Function $\CHECKBAL (\MANAGERS, \PUK, \NTEZ, \MTEZ)$ checks whether the balance of the account that associates with the public key $\PUK$ is greater or equal to $\NTEZ$ plus $\MTEZ$.
    \begin{itemize}
    \item[] $\MANAGERS(\PUK)$ = $\Angle{\BAL, \COU}$ and $\BAL \ge \NTEZ + \MTEZ$
    \end{itemize}
    \item Function $\CHECKCOU (\MANAGERS, \PUK)$ checks whether the current counter of an account that associates with the public key $\PUK$ is unlocked (its flag is False).
    \begin{itemize}
    \item[] $\MANAGERS(\PUK)$ = $\Angle{\BAL, \COU}$ and $\COU.b = \FALSE$
    \end{itemize}
   % \item Function \CHECKEXIST (\OPH, \PENDING, \ACCEPTED)\ checks whether an operation  \OPH\ exists in either \PENDING\ or \ACCEPTED.
    \item Function $\UPDATESUCC(\MANAGERS, \PUK, \PUK', \NTEZ, \MTEZ)$ updates the balance and the counter of the account that associates with the public key $\PUK$ and the balance of the account that associates with the public key $\PUK'$, where
    \begin{itemize}
        \item[]  $\PUK \mapsto \Angle{\BAL, (\hat\NTEZ, \TRUE)}$ $\Rightarrow$ $ \PUK \mapsto \Angle{\BAL - \NTEZ - \MTEZ, (\hat\NTEZ + 1, \FALSE)}$     
        \item[] $\PUK' \mapsto \langle  \BAL', \COU' \rangle$ $\Rightarrow$ $\PUK' \mapsto \langle \BAL' + \NTEZ, \COU' \rangle $         
    \end{itemize}

    \item Function $\UPDATECOU(\MANAGERS, \PUK, \BOOLEAN)$ updates the counter flag (True $=$ locked, False $=$ unlocked) of the account that associates with the public key $\PUK$, where
    \begin{itemize}
        \item[]  $ \PUK \mapsto \langle \BAL, (\NTEZ, \hat\BOOLEAN) \rangle $ $\Rightarrow$ $\PUK \mapsto \langle \BAL, (\NTEZ, \BOOLEAN) \rangle $ 
    \end{itemize}

    \item Function  $\UPDATECONSTR (\CONTRACTORS, \PUH, \NTEZ, \PARAMETER)$ updates the contractor that associates with the public hash $\PUH$, where
    \begin{itemize}
        \item[] $\PUH  \mapsto \langle   \CODE, \TIME, \BAL, \STORAGE, \rangle $ $\Rightarrow$ $\PUH  \mapsto \langle  \BAL + \NTEZ, \CODE, \\ \UPDATESTORAGE(\STORAGE,\CODE,\PARAMETER) \rangle $
    \end{itemize}

     \item Function $\UPDATESTORAGE(\STORAGE, \CODE,  \PARAMETER)$ returns the new storage by running the code $\CODE$ on the  current storage  $\STORAGE$ with  the input parameter $\PARAMETER$.

    \item Function $\CHECKPRG (\CODE)$ checks whether the code $\CODE$ is well type.

    \begin{itemize}
        \item[] $\CHECKPRG (\CODE) = \TRUE$ iff $\CODE : \TPAIR\ \TYPE_p\ \TYPE_s$
    \end{itemize}
    \item Function $\CHECKINIT (\CODE, \INIT)$ checks whether the initil stogare valude is well type with corresponding to the code $\CODE$.
    \begin{itemize}
        \item[] $\CHECKINIT (\CODE, \INIT) = \TRUE$ iff $\CODE : \TPAIR\ \TYPE_p\ \TYPE_s$ and $\STRING : \TYPE_s$
    \end{itemize}

    \item Function $\CHECKARG (\CONTRACTORS, \PUH, \PARAMETER)$ checks whether the input parameter  $\PARAMETER$ to the smart contract $\PUH$ in $\CONTRACTORS$ is well type with corresponding to its code.
    \begin{itemize}
        \item[] $\CHECKARG (\CONTRACTORS, \PUH, \PARAMETER) = \TRUE$ iff $\Delta (\PUH) =
  \TPAIR\ \TYPE_p\ \TYPE_s$ and $\PARAMETER : \TYPE_p$
    \end{itemize}

   % \item Function \CHECKGAS(\CODE, \INIT, \MTEZ) checks whether the gas that is consumed to run the code \CODE\ for the initial value \INIT\ is less or equal to \MTEZ.
    \item Function  $\CHECKGAS (\MANAGERS, \PUK,\NTEZ, \MTEZ)$/ $\CHECKGAS (\CODE, \INIT, \NTEZ, \MTEZ)$ / $\CHECKGAS (\CONTRACTORS, \PUH, \PARAMETER, \MTEZ)$ check whether the fee that is consumed to emit the opertion is less or equal to the fee \MTEZ. 
    %run the code \CODE\ that associates with the public hash \PUH\ for the paremeter \PARAMETER\ is less or equal to \MTEZ.
    \item Function \GENERATEOPH (\PUK, {\ADDR /\CODE},{\INIT /\PARAMETER}, \NTEZ, \MTEZ,  \TIME) generates an operation hash.

    \item Function \GENERATEHASH (\CODE, \TIME)  generates the public hash of a smart contract.

\end{enumerate}
\end{document}