### 2021-03-10 assertions/doc/thesis-bernhardt/assertion_highlighting.tex
Introduction first draft

\usepackage{listings, xcolor}

\definecolor{verylightgray}{rgb}{.97,.97,.97}

\lstdefinelanguage{Assertion}{
	keywords=[1]{entrypoint, assert, if, else, true, false, forall, exists},
	keywordstyle=[1]\color{blue}\bfseries,
	keywords=[2]{address, unit, never, bool, int, nat, string, chain_id, bytes, mutez, key_hash, key, signature, timestamp,
	option, or, pair, list, set, operation, lambda, map, big_map},	% types; money and time units
	keywordstyle=[2]\color{teal}\bfseries,
	keywords=[3]{left, right, some, none, cons, nil, _},
	keywordstyle=[3]\color{violet}\bfseries,
	identifierstyle=\color{black},
	sensitive=false,
	comment=[l]{//},
	morecomment=[s]{(*}{*)},
	commentstyle=\color{gray}\ttfamily,
	stringstyle=\color{red}\ttfamily,
	morestring=[b]',
	morestring=[b]"
}