### 2021-03-10 assertions/doc/thesis-bernhardt/michelson_highlighting.tex

\usepackage{listings, xcolor}

\definecolor{verylightgray}{rgb}{.97,.97,.97}

\lstdefinelanguage{Michelson}{
	keywords=[1]{parameter, storage, code},
	keywordstyle=[1]\color{blue}\bfseries,
	keywords=[2]{address, unit, never, bool, int, nat, string, chain_id, bytes, mutez, key_hash, key, signature, timestamp,
	option, or, pair, list, set, operation, ticket, lambda, map, big_map, sapling_transaction, sapling_state, bls12_381_g1,
	bls12_381_g2, bls12_381_fr},	% types; money and time units
	keywordstyle=[2]\color{teal}\bfseries,
	identifierstyle=\color{black},
	sensitive=false,
	comment=[l]{//},
	morecomment=[s]{(*}{*)},
	commentstyle=\color{gray}\ttfamily,
	stringstyle=\color{red}\ttfamily,
	morestring=[b]',
	morestring=[b]"
}

\lstset{
	language=Michelson,
	backgroundcolor=\color{verylightgray},
	extendedchars=true,
	basicstyle=\footnotesize\ttfamily,
	showstringspaces=false,
	showspaces=false,
	numbers=left,
	numberstyle=\footnotesize,
	numbersep=9pt,
	tabsize=2,
	breaklines=true,
	showtabs=false,
	captionpos=b
}