### 2021-03-18 assertions/doc/thesis-bernhardt/chapters/3-offline_design_tezos.tex

\chapter{Approach}\label{chap:approach}

With the specifications all set, the implementation section of this chapter assesses which of the necessary operations Tezos' smart contract language already provides, which extensions are needed and how the assertions can be made executable by Tezos' virtual machine.