### 2021-05-21 papers/contract-modules/fmbc2021/macros.tex

\usepackage{graphics}
\usepackage{listings}

\lstdefinelanguage{michelson}{
%  basicstyle=\ttfamily,
  backgroundcolor=\color{pink!20}, 
  classoffset=0,
  morekeywords={parameter,storage,code},
  keywordstyle=\bfseries, 
  % predefined types
  classoffset=1,
  morekeywords={nat,int,unit,string,or,pair,option,address,mutez,bool},
  keywordstyle=\color{blue},
  classoffset=2,
  morekeywords={CAR,CDR,CONS,DUP,EQ,FAILWITH,IF,IF_LEFT,LE,LEFT,NIL,DROP,PUSH,RIGHT,SENDER,SUB},
  keywordstyle=\color{magenta},
  sensitive=false,
  morecomment=[l]{\#},
  morecomment=[s]{/*}{*/},
  morestring=[b]",
}

\definecolor{pypypy}{rgb}{0.4,0.85,0.7}

\lstset{
  language=Caml,
  basicstyle=\ttfamily,
  backgroundcolor=\color{pypypy},
  morekeywords={requires,sender,amount,balance,storage,hash,module,contract,sig,val,paid,entrypoint,raises},
  classoffset=1,
  morekeywords={nat,int,unit,string,or,pair,option,address,bool},
  morekeywords={Tezos,pukh,tz,status,monad},
  keywordstyle=\color{blue},
}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "fmbc-main"
%%% End: