### 2022-02-16 papers/assertion-contract/appendix.tex
Add files via upload

\appendix
\chapter{Assertion Grammar in EBNF}\label{apx:grammar}
The following two versions of a grammar in extended Backus-Naur form (EBNF) implement the prefix and infix notations for the assertion syntax.

\section{Prefix Notation}
\lstinputlisting[basicstyle=\linespread{1.0}\fontfamily{lmr}\selectfont\small,
				 backgroundcolor=\color{cverbbg},
				 linewidth=14cm,
				 xleftmargin=0.5cm,
				 frame=lr,
				 framesep=8pt,
				 framerule=0pt,
				 captionpos=b,
				 numbers=none,
				 language=,
				 caption=Assertion grammar with prefix notation]{assertion_grammar_prefix.txt}

\section{Infix Notation}
\lstinputlisting[basicstyle=\linespread{1.0}\fontfamily{lmr}\selectfont\small,
				 backgroundcolor=\color{cverbbg},
				 linewidth=14cm,
				 xleftmargin=0.5cm,
				 frame=lr,
				 framesep=8pt,
				 framerule=0pt,
				 captionpos=b,
				 numbers=none,
				 language=,
				 caption=Assertion grammar with infix notation]{assertion_grammar_infix.txt}