### 2024-01-03 reports/2023-12-31/report.tex
end-of-year-report-dec-31-2023

\documentclass[a4paper,11pt]{article}

\usepackage[utf8]{inputenc}
\usepackage{url}
\usepackage{hyperref}
\usepackage{authblk}

\title{A Statically Checked Interface and a Checker for Tezos Smart Contracts}
\author{Peter Thiemann}
\author{Thi Thu Ha Doan}
\affil{University of Freiburg, Germany}
\begin{document}
\maketitle{}

\section{Progress Report July 2023 - December 2023}
\label{sec:executive-summary}

We proposed the following activities and deliverables for the
project. We started all planned activities for the reporting period
and we already started some planned activities for the next period.

Currently, the project is slightly delayed but making good progress. Ms.\ Doan
returned from  maternal leave in October 2023.

\subsection{Proposed activities for January 2023 - June 2023}

\begin{itemize}
\item Activity: Development of a dynamic logic as a formal foundation for symbolic interpretation of Michelson. \\
  $\Rightarrow$ This activity progressed significantly.  The master's
  thesis has been completed in November and is currently under review (by the examiner).
\item Deliverable: Formalization (in Agda) of a dynamic logic for a core fragment of Michelson with soundness proof. \\
  $\Rightarrow$ This deliverable is available  at
  \url{https://github.com/Barnaboss/thesis-michelsonDL}. It will be
  released after the thesis review period.
\item Deliverable: Paper about this formalization (writeup starting
  July 2023). \\
  $\Rightarrow$ The paper will be based on the master's thesis, so the final writing
  starts after completion of the thesis review. The goal is to submit to a
  conference during spring.
\end{itemize}

\subsection{Proposed activities for July 2023 - December 2023}
\label{sec:july-2023-december}

\begin{itemize}
\item Activity: Implementation of symbolic interpreter for a core
  fragment of Michelson in OCaml based on the SMT solver Z3. \\
  $\Rightarrow$ This activity has been driven by Doan between January and June
  2023 and then further after her return from maternity leave. 
\item Deliverable: Symbolic interpreter. \\
  $\Rightarrow$ The current prototype of the symbolic interpreter is available here
  \url{https://github.com/proglang/tezos-project/tree/master/static_checker}
\item Deliverable: Doan completed two case studies with example contracts.
\item Deliverable: Paper or tutorial on this tool (writeup started,
  submission planned for a deadline in spring) \\
  $\Rightarrow$ A draft is available here \url{https://github.com/proglang/tezos-project/tree/master/symbolic-execution/paper}
\end{itemize}

\subsection{Proposed activities for January 2024 - June 2024}
\label{sec:january-2024-june}

Not started.

\begin{itemize}
\item Activity: Design and implementation of a DSL for static contract properties. This DSL will compile to SMT and interface with the symbolic interpreter to perform verification.  
\item Deliverable: Implementation of the DSL aka contract checker.
\item Deliverable: Report on this tool.
\end{itemize}

\subsection{Properties activities for July 2024 - December 2024}
\label{sec:july-2024-december}

Not started.

\begin{itemize}
\item Activity: Complete design and implementation of the contract module DSL including contract orchestration. 
\item Deliverable: Full implementation of contract modules.
\item Deliverable: Followup on our FMBC paper. 
\end{itemize}

\bibliography{report}
\bibliographystyle{plainurl}
\end{document}