### 2024-06-20 reports/2024-06-30/report.tex
draft report

\documentclass[a4paper,11pt]{article}

\usepackage[utf8]{inputenc}
\usepackage{url}
\usepackage{hyperref}
\usepackage{authblk}

\title{A Statically Checked Interface and a Checker for Tezos Smart Contracts}
\author{Peter Thiemann}
\author{Thi Thu Ha Doan}
\affil{University of Freiburg, Germany}
\begin{document}
\maketitle{}

\section{Progress Report January 2024 - June 2024}
\label{sec:executive-summary}

We proposed the following activities and deliverables for the
project. We started all planned activities for the reporting period
and we already started some planned activities for the next period.

Currently, the project is still behind schedule, but making good
progress after Ms.\ Doan's return from  maternal leave in October 2023.

\subsection{Proposed activities for January 2023 - June 2023}

\begin{itemize}
\item Activity: Development of a dynamic logic as a formal foundation for symbolic interpretation of Michelson. \\
  $\Rightarrow$  The master's thesis has been completed in November and grading is complete.
\item Deliverable: Formalization (in Agda) of a dynamic logic for a core fragment of Michelson with soundness proof. \\
  $\Rightarrow$ This deliverable is available  at
  \url{https://github.com/Tezos-Project-Uni-Freiburg/michelson-dynamic-logic}. It will be
  released as an artifact at the time the paper is published.
\item Deliverable: Paper about this formalization (writeup starting
  July 2023). \\
  $\Rightarrow$ The paper is based on the master's thesis. The Agda code
  from the thesis as well as the text had to undergo significant
  revision. The paper was submitted to ECOOP on April 18 and has been
  conditionally accepted. The revision is due July 7.
\end{itemize}

\subsection{Proposed activities for July 2023 - December 2023}
\label{sec:july-2023-december}

\begin{itemize}
\item Activity: Implementation of symbolic interpreter for a core
  fragment of Michelson in OCaml based on the SMT solver Z3. \\
  $\Rightarrow$ This activity has been driven by Doan between January and June
  2023 and then further after her return from maternity leave. 
\item Deliverable: Symbolic interpreter. \\
  $\Rightarrow$ The current prototype of the symbolic interpreter is available here
  \url{https://github.com/proglang/tezos-project/tree/master/static_checker}
\item Deliverable: Doan completed two case studies with example contracts.
\item Deliverable: Paper or tutorial on this tool (writeup finished,
  first submission January 18, second submission May 31) \\
  $\Rightarrow$ Current submission is available here \url{https://github.com/proglang/tezos-project/tree/master/symbolic-execution/paper}
\end{itemize}

\subsection{Proposed activities for January 2024 - June 2024}
\label{sec:january-2024-june}

Not started.

\begin{itemize}
\item Activity: Design and implementation of a DSL for static contract properties. This DSL will compile to SMT and interface with the symbolic interpreter to perform verification.  
\item Deliverable: Implementation of the DSL aka contract checker.
\item Deliverable: Report on this tool.
\end{itemize}

\subsection{Properties activities for July 2024 - December 2024}
\label{sec:july-2024-december}

Not started.

\begin{itemize}
\item Activity: Complete design and implementation of the contract module DSL including contract orchestration. 
\item Deliverable: Full implementation of contract modules.
\item Deliverable: Followup on our FMBC paper. 
\end{itemize}

\bibliography{report}
\bibliographystyle{plainurl}
\end{document}